---
title: "1808-dsc351-351m-451-homework6-ISLR-ModVarSelect"
author: "Roger H. French, JiQi Liu"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

Use ISLR Chapter 6, Linear Model Selection & Regularization
And ISLR Chapter 7, Moving Beyond Linearity for this assignment

1 Questions, 5 points total.

# ISLR Exercise 6.10  (6 parts a-f, 5 points total)

We have seen that as the number of features used in a model increases,

  - the training error will necessarily decrease, 
    - but the test error may not.

We will now explore this in a simulated data set.

## (a) 

A data set (HW6-testdata.csv) has been generated 

  - with p = 10 features, 
  - n = 1,000 observations, 
  - and an associated quantitative response vector 
    - generated according to the model

$$ Y = X \beta + \epsilon $$
where $\beta$ has some elements that are exactly equal to zero.

The $\beta$ values 

  - are randomly generated coefficients 
  - which are with the p features (X1-10) 
    - to produce the y response.

Read in and describe the data set. 

```{r}
# Put your code here, with comments and good style and syntax
```

## (b) 

Split your data set 

  - into a training set containing 100 observations
  - and a test set containing 900 observations 
    - using the caret package.

```{r}
library(caret)
# Put your code here, with comments and good style and syntax
```

## (c) 

Perform best subset selection on the training set,

  - and plot the training set MSE 
  - associated with the best model of each size.

```{r}
# Put your code here, with comments and good style and syntax
```

## (d) 

Plot the test set MSE 

  - associated with the best model of each size. 

```{r}
# Put your code here, with comments and good style and syntax
```

## (e) 

For which model size 

  - does the test set MSE 
  - take on its minimum value? 

Comment on your results. 

If it takes on its minimum value 

  - for a model containing only an intercept 
    - or a model containing all of the features, 
  - then play around with the way that you are generating the data in (a) 
    - until you come up with a scenario 
    - in which the test set MSE is minimized 
    - for an intermediate model size.

```{r}
# Put your code here, with comments and good style and syntax
```

## (f) 

In this data set, 

  - 3 of the $\beta$ coefficients were made to be 0, 
  = how is this reflected in your test MSE fit? 

Comment on the coefficient values 

  - what are their signs and magnitudes?

```{r}
# Put your code here, with comments and good style and syntax
```


# Cites

* Gareth James, Daniela Witten, Trevor Hastie, and Robert Tibshirani. An Introduction to Statistical Learning: With Applications in R. 1st ed. 2013, Corr. 5th printing 2015 edition. Springer Texts in Statistics. New York: Springer, 2013.

