---
title: 'CWRU DSCI351-451: Project 4 - SamSensor Supervised Machine Learning'
author: "Roger H. French, JiQi Liu"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
  html_document:
    font-size: 10em
    self_contained: yes
    toc: yes
urlcolor: blue
---

<!--
# Script Name: 1711-dsci351-451-project4.Rmd
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->


\setcounter{section}{4}
\setcounter{subsection}{1}
\setcounter{subsubsection}{0}

#### Semester Project 4: SamSensor Supervised Machine Learning (ML)

[Machine learning](https://en.wikipedia.org/wiki/Portal:Machine_learning) 

  - is a powerful tool for data analysis, 
  - allowing for modeling and prediction 
    - of complex data structures.
  
In this lab we will be using 

  - the rotational and acceleration data 
    - from Samsung Galaxy S3 cell phones
    - from the [UCI Machine Learning Repository](https://archive.ics.uci.edu/ml/datasets/human+activity+recognition+using+smartphones)
    - a [YouTube about the data is here](https://www.youtube.com/watch?v=XOEN9W05_4A)
  - to predict what activity people holding the phones are doing.
  
This data contains all of the variables from the phone 

  - and the "activity" column that shows 
    - what activity was being done at each measurement. 
    
The features and features_info files 

  - can help you understand the variables.
  
We will be focusing on 3 machine learning methods, 
  
  - [Regression (or Decision) Tree](https://en.wikipedia.org/wiki/Decision_tree_learning), 
    - using the rpart package
  - [Random Forest](https://en.wikipedia.org/wiki/Random_forest),  
    -  using the randomForest package
  - [Support Vector Machine](https://en.wikipedia.org/wiki/Support_vector_machine),
    - using the e1071 package.

----------------------

##### Question 1: Supervised and Unsupervised Machine Learning

Explain supervised and unsupervised machine learning in your own words.

ANSWER:

Read in the SamsungData.rda data for this lab, 

  - how many predictors and responses are there, 
  - how many different activities are there 
  - and what are they?

You will need to split this data into training and testing datasets.

  - use the caret package to do this splitting
  - its a swiss army knife for accessing > 100 ML methods

```{r}
library(caret)
 ??caret::caret

```
ANSWER:

In your opinion, 

  - is variable selection necessary for this data set 
  - and if so can it be reasonably done by a human inferring 
    - or assuming which variables are important?

ANSWER:

- Explain why this lab is supervised machine learning and not unsupervised

ANSWER:


----------------------

##### Question 2: Data Cleaning and "-" in Column Names

There have been some changes to R recently 

  - and "-" can no longer be used in column names 
  - without causing problems so we have to do some data cleaning

Such is life for data scientists

  - Edit the column names to format them better
  - Replace all parentheses  and commas "(" ")" and "," with periods "." 
  - Replace all dashes "-" with underscores "_"
  - Regular expression (regex) can help you do this

```{r}

```

Lets build a simple Regression Tree 

  - to try to predict the activity using the training data
  
Use the rpart() function from the rpart package 

  - to build a basic regression tree, 
  - be sure and keep this model, 
  - we'll use it later

```{r}
library(rpart)
?rpart

```

Plot the model with plot() and text() 

  - and plot the error per split with the rsq.rpart() function
    - the splits in the trees show the level of branches in the tree

```{r}

```

How many splits were there in your regresion tree and 

  - does this make sense based on the number of activities?

ANSWER:

Explain how someone could estimate the activity using this tree 

  - if they had unlabled data

ANSWER:


----------------------

##### Question 3: Regression Trees and Random Forest ML

In your own words, 

  - explain what a random forest is and 
  - what is its relationship with regression trees 
    - like the one you just fitted in Question 2?

ANSWER:

Use the randomForest function 

  - in the randomForest package 
  - to fit a model to the training data

```{r}
library(randomForest)
?randomForest

```

Show your results with the plot() and print() functions on the model, 

  - explain the meaning of result.

```{r}

```

Use the importance() function 

  - to find the most important variable as determined by the random forest
  - what was this variable measuring?

Using ggplot and geom_density, 

  - plot the distribution of this variable 
    - for each activity in one plot, 
  - can you explain why this variable is important in classification?

```{r}


```

Were there any activities 

  - that the model had more trouble distinguishing, 
  - if so does is make sense that they were hard to distinguish?

ANSWER:

What was the error of the random forest 

  - and how did it compare to the regression tree?

ANSWER:


----------------------

##### Question 4: Support Vector Machine ML Model

In your own words 

  - explain what a support vector machine is.

ANSWER:

Use the e1071 package 

  - [I don't know why they named it that either][1] 
  - to build an SVM model on the data set

```{r}
library(e1071)
?e1071

```

----------------------

##### Question 5: Training and Testing Your 3 ML Models

Read in the testing data and 

  - use the predict() function 
    - to predict the activity of the test data with
      - your regression tree, 
      - your random forest, 
      - and your SVM model

```{r}

```

Compare the predictions for each activity and 

  - compare them to the labeled activity in the testing data
    - What is the error of each model?

```{r}

```

Which model do you think 

  - does the best job of predicting the activity?
  - explain your decision

ANSWER:


----------------------

#### Links

[1]: The package authors belonged to vienna university of technology [at that time]. 
[Institut für Statistik und Wahrscheinlichkeitstheorie](https://tiss.tuwien.ac.at/adressbuch/adressbuch/orgeinheit/1620). Their statistic department has code :e107. And they were in the computational intelligence section of that department, called e1071. 

http://www.r-project.org 

http://rmarkdown.rstudio.com/  

<!--
# Keep a complete change log history at bottom of file.
# Complete Change Log History
# v0.00.00 - 1405-07 - Nick Wheeler made the blank script
##########

