---
title: "CWRU DSCI351-451: Dataframe 7 Ways and Violent Crime"
author: "Roger H. French, JiQi Liu"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 \setcounter{section}{0}
 \setcounter{subsection}{1}
 \setcounter{subsubsection}{0}
 
 <!-- 
 How to make comments inside Rmarkdown
# Script Name: My class notes template for Fall 2016
# Purpose: This is a template Rmd file to start a new class from
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - Filename.Rmd - Roger French started this blank Rmd script
-->

<!-- Or on a single line like this -->
 


#### First: Guess The Correlation Game

[Guess The Correlation Game](http://guessthecorrelation.com/)


#### Examine a Data Frame in R with 7 Basic Functions

When one first started learning R, 

  - it seems way more complicated than what you were used to 
  - with looking at spreadsheets in Microsoft Excel. 
  
When I started working with data frames in R, 

  - it didn't seem quite as easy to know what I was looking at.

I've since come to see the light. 

  - While there is a bit of a learning curve to get a handle on it, 
  - viewing data in R is infinitely more flexible than doing so in Excel. 

In this discussion, I'll cover the most basic R functions 

  - for examining a data set and 
  - explain why they're important.

Understanding how to get a simple overview of the data set 

  - has become a huge time saver for me. 
  - If you aren't familiar with these functions, you need to be. 
  - If you're anything like me, 
    - you'll use them first for every single data set you consider.

All of the functions I'm discussing here come in the base R Utils package, 

  - so there's no need to install any additional packages. 
  
Here are the functions, with links to their documentation:

  - dim(): shows the dimensions of the data frame by row and column
  - str(): shows the structure of the data frame
  - summary(): provides summary statistics on the columns of the data frame
  - colnames(): shows the name of each column in the data frame
  - head(): shows the first 6 rows of the data frame
  - tail(): shows the last 6 rows of the data frame
  - View(): shows a spreadsheet-like display of the entire data frame

##### Now, let's import a data set see how each of these functions works. 

First, here's the code:

### Import a data set on violent crime by state and assign it to the data frame "crime"

```{r}
crime <- read.csv("http://vincentarelbundock.github.io/Rdatasets/csv/datasets/USArrests.csv", stringsAsFactors = FALSE)
```

Now, let's take a look at the output, 

  - so we can see what happens when the code is run.

First, we'll look at the 

  - dim(), 
  - str(), 
  - summary(), and 
  - colnames()  functions:

```{r}
### Call the functions on crime to examine the data frame
dim(crime)
```

  - dim(): In the crime data set, we can see immediately that 
    - there are only 50 rows and 5 columns. 
    - This function is useful, because it tells us whether it would be okay to print the entire data frame to the console. With this data set, it's probably okay. 
    - If, however, there were 5,000 rows and 50 columns, we'd definitely want to view the data frame in smaller chunks.

```{r}
### Call the functions on crime to examine the data frame
str(crime)
```

  - str(): The structure of the crime data set also 
    - tells us the number of rows (observations) and columns (variables), 
    - but it provides even more information. 
      - It tells us the column names, 
      - the class of each column (what kind of data is stored in it), 
      - and the first few observations of each variable.

```{r}
### Call the functions on crime to examine the data frame
summary(crime)
```

  - summary(): The summary provides descriptive statistics including 
    - the min, max, mean, median, and quartiles of each column. 
    - For example, we can see in the crime data set that 
      - the average murder rate across all states is 7.8 for every 100k people.

```{r}
### Call the functions on crime to examine the data frame
colnames(crime)
```

  - colnames(): This function prints a vector of the column names, 
    - which can be useful if you're trying to reference a particular column. 
    - For the crime data set, we can see that the state column has no name. 
    - Knowing this, we may want to assign it a name before going forward in our analysis.


Now, let's take a look at the head() and tail() functions:

```{r}
### The head() and tail() functions default to 6 rows, but we can adjust the number of rows using the "n = " argument
head(crime, n = 10)
tail(crime, n = 5)
```

  - head(): This function defaults to printing the first 6 rows, 
    -but we've decided to call the first 10. 
    - In the crime data set, this gives us the data on states Alabama through Georgia.
  - tail(): The same as head(), 
    - except this function prints the end of the data frame. 
    - In this case, we've called the last 5 observations, 
      - so we can see the data on Virginia through Wyoming.

Finally, let's take a look at the window that appears when we call the View() function:

```{r}
### While the first 6 functions are printed to the console, the View() function opens a table in another window
View(crime)
```

  - View(): This window provides vertical and horizontal 
    - (if enough columns to justify) 
    - scroll bars for you to browse the entire data set. 
    - It looks exactly like an Excel spreadsheet' 
      - you just can't manipulate any of the data. 
      - (Note: make sure you use a capital 'V' when calling this function; it's case sensitive).

That's it! Getting comfortable with these functions should make it easier for you to work with data frames in a more logical and efficient manner. 


#### Cites

  - Douglas E Rice [https://www.r-bloggers.com/examine-a-data-frame-in-r-with-7-basic-functions/amp/](https://www.r-bloggers.com/examine-a-data-frame-in-r-with-7-basic-functions/amp/)
  