---
title: 'CWRU DSCI351-351M-451: Exploratory Data Science'
author: "Prof.:Roger French, TA:Alan Curran"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

 \setcounter{section}{1}
 \setcounter{subsection}{2}
 \setcounter{subsubsection}{0}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 6, fig.height = 4.5) 
```
 
 <!-- 
 How to make comments inside Rmarkdown
-->

<!-- Or on a single line like this -->
 
 
### R Learning Resources

  * Peng: R Programming for Data Science (Book, in readings)
  * Roger Peng's [Youtube Playlist for 4 weeks of Coursera R Programming](https://www.youtube.com/playlist?list=PLjTlxb-wKvXNSDfcKPFH2gzHGyjpeCZmJ)
  - [Lynda.com R Courses](https://www.lynda.com/R-training-tutorials/1570-0.html) through CWRU
  
  
### SDLE Teatime Learning

  - 2016 year was intro to datascience, R, Python, Git, LaTeX
  - 2017 was more advanced topics including Hadoop and Spark and SparklyR
  - 2018 continued with more advanced topics and review
  - [2016 SDLE Teatime Repo](https://bitbucket.org/cwrudsci/16-sdle-teatime) 
  - [2017SDLE Teatime Repo](https://bitbucket.org/cwrudsci/17-sdle-teatime) 
  - [2018 SDLE Teatime Repo](https://bitbucket.org/cwrudsci/18-sdle-teatime/)
  - [SDLE Teatime Youtube Playlist](https://www.youtube.com/channel/UCer8o-Bxh-xef_6Kdmq4Kvg)
 
```{r,echo=FALSE,message=FALSE}
options("digits" = 5)
options("digits.secs" = 3)
```

#### What we need to do this week

  1. Setup VDI
   - Rstudio
   - Drag icons of R, Rstudio, Git Bash, Spyder, Jupyter Notebook, HipChat to desktop 
  2. Setup Git
    - make H:\\Git folder
    - git config name and email
  3. Setup Twitter
  4. Setup StackExchange
  5. Setup Slack
  6. Setup Kaggle
  8. Git Clone
    - 18-sdle-tea-time
      - for quick introduction to data science techniques and tools
    - For Class-Prof Repo
      - Clone your forked Class Repo
      
 
### Bash: The language of the Linux Console
 
  * Bash is the command line processor of the Linux Console
  * R has its own command line processor for the R Console
  * Bash is the default Console for both Linux and for Mac
    + Mac's are based on BSD-Unix OS
    + A close variant of Linux, only different by the licensing
  * Windows uses the DOS command line processor in its 'Command Prompt'

 
#### On our ODS VDI's we use Git Bash to work with Git

  * MinGW64 is a little Linux OS running inside Windows
    + It has the standard Bash commands
    + And tools like vim (the visual text editor)
    
#### Lets see some Bash Commands we'll be using

  * ls is the "list" command, to get a directory of files and folders
  * pwd is the "present working directory" command, to know where you are
  * cd is "change directory" 
  * .. refers to the directory one up from where you are
  * so "cd .." moves you up one directory
  * and "cd Downloads" would move you down into Downloads directory (if it exists)
  * To copy a file use "cp"
  * To move a file use "mv"  [Look up linux command syntax]
  * To make a new directory use "mkdir"
    
#### A good resource for Bash Commands and Man pages

  * Is [An A-Z Index of the Bash command line for Linux](http://ss64.com/bash/)
  * There are many other resources too

### Now lets start working with our local Git Server

  * Using Git Bash to talk to it
    * Git is also a linux program 
  * All Git commands are entered at the Bash Prompt
  * All Git commands start with "git "
 ` * So that the Bash prompt know who to send the subsequent command to
    
##### Check your Git Server Configuration
          
          git config --list

##### Essential git config --global's, Set your user info

  - git config --global user.name "[name]"
  - git config --global user.email "[email address]"
  - git config --global color.ui auto

#### First we need to go up to Bitbucket and "Fork" the Prof. Repo

  * This will give you a copy of Prof. Repo
    * In your personal account area
    * You want to change the ending from "Prof" to your caseID
    
#### Now you want to open Git Bash on Windows

  * You need to save your Repos on your H: drive, NOT C drive
    * C Drive is restricted
    * H Drive is your personal area that follows your caseID login
  * So in Git Bash
    * "pwd"" will tell you your present working directory
    * "cd .." moves up a directory
    * "pwd" to see where you have moved
    * Now change to H: "cd /h"
    * "pwd" see where you are
    * "ls" see what files are there
    * "mkdir Git"  this will make a new directory at H:Git
      + So you'll keep all your repositories under H:Git

#### Important Note: Windows ignores case, Linux and BSD-Unix (Mac) respect case

  * So Git and git are the same on windows for a folder
  * They are totally different on Linux or Mac
  * Best practice Use capitals sparingly
  * About only useful place is in CamelBack filenames
    * Since I said, no spaces in filenames
    * To make things readable, you can do CamelBack
    * Example: 1608DSCI351-ThisIsMyReport.rmd
    
### Now lets Clone your personal class repo    
    
  * Now you want to Clone your personal class repo
    * This is a one time operation
    * To copy all the files and folders down to your local computer
  * In Git Bash, you want to be at H:Git or h:Git
    Check with pwd
  * Now go to your personal class repo on Bitbucket
    * And find the clone command
    * For windows choose https protocol (Not ssh)
  * Copy the command
    * Its something like this 
    * "git clone https://vuvlab@bitbucket.org/cwrudsci/18f-dsci351-351m-451-CASEID.git"
  * Now that that is on your clipboard
  * Go to you Git Bash, and use "Shift-Insert" (Not cntrl-v)
    * To copy it onto the Bash Command line
    * Hit enter, and watch a full copy of your repo being copied locally
    
### Now lets pull and push changes from to your repo

  * cd into your repo's top folder
    * This can be donw with tab completion
    * "cd 16f" and hit tab, it auto completes
  * Now type "git pull" To see if there are any changes up on bitbucket
    And to pull these down and merge them in

### Making local changes, Adding, Commmitting and Pushing

  * Now change a local file by adding something into it
  * Now you add this changed file to be tracked by Git
    * "git add ." or "git add --all"
  * Now commit your changes
    * "git commit -m 'I have changed the readme.md' "
  * Now push your changes up to Bitbucket, to your personal repo
    * "git push"
    
### Links
 
  - http://www.r-project.org 


