---
title: "hyperSpec Package Review"
author: "Devin Gordon"
date: "Tuesday, October 30, 2018"
output: ioslides_presentation
---

<!--
# Script Name: hyperSpecReview.Rmd
# Purpose: This is an R Markdown to review the utilities of the hyperSpec R package
# Authors: Devin A. Gordon
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - Devin Gordon started blank script
# v0.00.02 - Presentation renamed to V2 and changed to html output

# Rmd code goes below the comment marker!
-->

#Background

## Spectra - What are they?

  - "a specific set of values that can vary infinitely within a continuum"
  - Generally: 2-dimensional (sometimes higher) dimensional data, 
    - where the first dimension identifies points along a continuum, 
    - and the second dimension the values of a response of interest 
    - at the corresponding points
  - Examples: Optical, Mass, Political Alignment

## Spectral Data - How do we store them?

  - Proprietary file format: not so useful
  - .csv files (ASCII): better ~ excel, R
  - .spc files (Binary): good ~ R

## hyperSpec - How to manipulate .spc files in R

  - http://hyperspec.r-forge.r-project.org/
  - Handles spectral data
  - Features: Convenient import, spectral range selection, shfiting,  plotting/viewing, normalization, smoothing, correction, arithmetic, initial data analysis


#hyperSpec Examples
```{r, echo=FALSE}
library(hyperSpec)
```

## File Import

```{r}
# Read one .spc file and check its structure
# setwd("V:/vuv-data/instr/spectra/")
curfile <- "./data/pet_unstab_cyclic_quv/sa19603_00_ex320.spc"
step0 <- read.spc(curfile)
class(step0)
step0
```

## Basic Plotting

```{r}
#Plot with hyperSpec base plotting

plotspc(step0)
```

## Multiple .spc Files?

- (Purposefully without a loop)
```{r}
# Read multiple .spc files

setwd("./data/corrected_pet_unstab_cyclic_quv_ex320/")
curfile <- "./sa19603_00_ex320.spc"
step0 <- read.spc(curfile)
curfile <- "./sa19603_22_ex320.spc"
step1 <- read.spc(curfile)
curfile <- "./sa19603_23_ex320.spc"
step2 <- read.spc(curfile)
curfile <- "./sa19603_24_ex320.spc"
step3 <- read.spc(curfile)
curfile <- "./sa19603_25_ex320.spc"
step4 <- read.spc(curfile)
curfile <- "./sa19603_26_ex320.spc"
step5 <- read.spc(curfile)
curfile <- "./sa19603_27_ex320.spc"
step6 <- read.spc(curfile) 
curfile <- "./sa19603_28_ex320.spc"
step7 <- read.spc(curfile)
```

## Combine into one object (Good News!)

```{r}
# Use Collpase to combine spectra

spec <- collapse(step0, step1, step2, step3, step4, step5, step6, step7)
class(spec)
spec
```

## Basic Plotting

```{r}
plotspc(spec)
```

## Spectral Range Selection

```{r}
#Subset and redifine the hyperSpec object according to wavelength

spec <- spec[,, min ~ 600]
spec
```

## Basic Plotting


```{r}
plotspc(spec)
```

## Spectral Arithmetic (Subtraction)
```{r}
#Perform spectral subtraction

sub <- step7 - step0
plotspc(sub[,, min ~ 600])
```

## Removing Bad Data
```{r}
# Let's look at some fake IR data
ir.spc <- chondro
ir.spc
plotspc(ir.spc, spc.nmax = length(ir.spc))
```

## Removing Bad Data
```{r}
#Define any point above 1500 as bad
high.int <- apply(ir.spc > 1500, 1, any)

#Maximum should atleast be 0.1 
low.int <- apply(ir.spc, 1, max) < 0.1

#Apply Conditions
ir.spc <- ir.spc[!high.int & !low.int]
```

## Removed the bad spectra
```{r}
plotspc(ir.spc, spc.nmax = length(ir.spc))
```

## Baseline Correction (One of many options)
```{r}
# Apply basline correction function
blcorr <- spc.fit.poly.below(ir.spc)

# Subtract away correction from original
ir.spc <- ir.spc - blcorr
```

## Corrected Spectra!
```{r}
plotspc(ir.spc)
```

# Example of use

## Load Cyclic QUV Absorbance Data Set

```{r}
# setwd("v:/vuv-data/proj/3M")

## Read keyfile(s) and set classes

pet.key <- read.csv("./data/pet_unstab_cyclic_quv/3m_sample_key.csv")
pet.key$Sample <- as.character(pet.key$Sample)
pet.key$Product <- as.character(pet.key$Product)
pet.key$Exposure <- as.character(pet.key$Exposure)
pet.key$Step_Retained <- as.character(pet.key$Step_Retained)

# Read step 1 CyclicQUV Optical Absorbance Data

filenames <- list.files(path = "./Cary/CyclicQUV/step1/", pattern = "\\.SPC$")
numfiles <- length(filenames)
n <- 0
cq.abs.step1 <- list(rep(NaN,numfiles))
for (i in filenames) { 
  filepath <- paste("./Cary/CyclicQUV/step1/",i,sep = "")
  n = n + 1
  cq.abs.step1[[n]] <- read.spc(filepath)
  
  # Add step data
  
  cq.abs.step1[[n]]@data$step <- 1
  
  # Match material info from key file and add to hyperSpec object
  
  ## Extract sample number from file name using sub (grep) to pull the sample number from the file name match to row in the keyfile
  
  temp <- sub("-es01-ms01-mn01_uvvs01.SPC","",i)
  samplenum <- sub("_",".",temp)   
  cq.abs.step1[[n]]@data$sample <- samplenum
  
  row <- pmatch(samplenum, pet.key[,1])                  
  ifelse(is.na(row) == 'FALSE',cq.abs.step1[[n]]@data$material <- pet.key[row,2], 0 )
}

cq.abs.step1 <- collapse(cq.abs.step1[1:numfiles])
cq.abs.step1 <- orderwl(cq.abs.step1)

# Read step 2 CyclicQUV Optical Absorbance Data

filenames <- list.files(path = "./Cary/CyclicQUV/step2/", pattern = "\\.SPC$")
numfiles <- length(filenames)
n <- 0
cq.abs.step2 <- list(rep(NaN,numfiles))
for (i in filenames) { 
  filepath <- paste("./Cary/CyclicQUV/step2/",i,sep = "")
  n = n + 1
  cq.abs.step2[[n]] <- read.spc(filepath)
  
  # Add step data
  
  cq.abs.step2[[n]]@data$step <- 2
  
  ## Extract sample number from file name using sub (grep) to pull the sample number from the file name match to row in the keyfile
  
  ## Extract sample number from file name
  
  temp <- sub("-es02-ms02-mn01_uvvs01.SPC","",i)
  samplenum <- sub("_",".",temp)   
  cq.abs.step2[[n]]@data$sample <- samplenum
  
  row <- pmatch(samplenum, pet.key[,1])                  
  ifelse(is.na(row) == 'FALSE',cq.abs.step2[[n]]@data$material <- pet.key[row,2], 0 )
}

cq.abs.step2 <- collapse(cq.abs.step2[1:numfiles])
cq.abs.step2 <- orderwl(cq.abs.step2)

# Read Baseline DATA

filenames <- list.files(path="./Cary/Baseline/", pattern = "\\.spc$")
numfiles <- length(filenames)
n <- 0
baseline <- list(rep(NaN,numfiles))

for (i in filenames) { 
  filepath <- paste("./Cary/Baseline/",i,sep ="")
  n = n+1
  baseline[[n]] <- read.spc(filepath)
  
  # Add step data (could be performed using a matching function as below for more extensibility)
  
  baseline[[n]]@data$step <- 0
  
  # Match material info from key file and add to hyperSpec object
  
  ## Extract sample number from file name using sub (grep) to pull the sample number from the file name match to row in the keyfile
  
  temp2 <- sub("./Cary/Baseline/","",i)
  temp <- sub("-es00-ms00-mn01_uvvs01.spc","",temp2)
  samplenum <- sub("_",".",temp)   
  baseline[[n]]@data$sample <- samplenum
  
  row <- pmatch(samplenum, pet.key[,1])                  
  ifelse(is.na(row) == 'FALSE',baseline[[n]]@data$material <- pet.key[row,2], 0 )
}

baseline <- collapse(baseline[1:6])
baseline <- orderwl(baseline)
```

Lets check out the contents of each hyperSpec Object

```{r}
baseline

cq.abs.step1

cq.abs.step2
```

We'll put these all together

```{r}
# Combine baselines with step 1

cq.abs <- collapse(baseline, cq.abs.step1, cq.abs.step2)

cq.abs

```

# Utilizing spectral arithmetric to change ordinate axis

```{r}
# Divide by sample thickness for Abs/cm

t.melinex243 <- 0.005
t.melinex238 <- 0.0127
t.melinex626 <- 0.0127
t.melinex618 <- 0.0127
t.teteronu2 <- 0.005
t.teteronu2l92w <- 0.0125

for (i in 1:length(cq.abs@data$material)) {
  
  if (cq.abs@data$material[i] == "B-Melinex243") {cq.abs[i] <- cq.abs[i]/t.melinex243}
  if (cq.abs@data$material[i] == "B-Melinex238") {cq.abs[i] <- cq.abs[i]/t.melinex238}
  if (cq.abs@data$material[i] == "A-Melinex626") {cq.abs[i] <- cq.abs[i]/t.melinex626}
  if (cq.abs@data$material[i] == "A-Melinex618") {cq.abs[i] <- cq.abs[i]/t.melinex618}
  if (cq.abs@data$material[i] == "B-TeteronU2") {cq.abs[i] <- cq.abs[i]/t.teteronu2}
  if (cq.abs@data$material[i] == "B-TeteronU2L92W") {cq.abs[i] <- cq.abs[i]/t.teteronu2l92w}
}

cq.abs

# Remove errenous points (negative)

`cq.abs` [[`cq.abs` < 0]] <- NA

cq.abs
```


## Better plotting using ggplot2 graphics

```{r, echo=FALSE}

# Plot directly using ggplot2 graphics

qplotspc (cq.abs[,, 304  ~ 800], mapping = aes (x = .wavelength, y = spc, color = as.factor(step), group = .rownames), spc.nmax = length(cq.abs)) +
  facet_grid (material ~ ., scales="free_y") + xlab ("Wavelength [nm]") + ylab("Abs/cm") +
  geom_line(size=1) +
  ggtitle(paste("Optical Absorbance Spectra - CyclicQUV")) +
  scale_x_continuous(breaks = seq(0, 800, by = 100)) +
  scale_colour_discrete(name  ="Exposure Time [hr]",
                        breaks=c("0", "1", "2"),
                        labels=c("0", "168", "336")) +
  theme(legend.title = element_text(size=20,face="bold"), 
        legend.text = element_text(size=20, face="bold"),
        axis.text=element_text(size=18,face="bold"),
        axis.title=element_text(size=30,face="bold"), 
        plot.title=element_text(size=30,face="bold"), strip.text.y = element_text(size = 12))
```

# Peak Information extraction

We can directly address the intensity value at a specific wavelength

```{r}
spec[2,,460]@data$spc
```

Using this concept, we can also find full width at half max by approximating the wavelengths at half the intensity of the peak

```{r}

lower <- approx(spec[2,,330 ~ 460]$spc, spec[2,,330 ~ 460]@wavelength, xout = c(spec[2,,460]$spc/2))
lower

upper <- approx(spec[2,,460 ~ 600]$spc, spec[2,,460 ~ 600]@wavelength, xout = c(spec[2,,460]$spc/2))
upper

fwhm <- upper$y - lower$y
fwhm

```

Imagine this as a function that could operate on multiple spectra

```{r}
fwhm.spc <- function(spec, peak.wl, lower.bound, upper.bound){

  peak.wl <- as.numeric(peak.wl)
  
   lower <- approx(spec[,,lower.bound ~ peak.wl]$spc, spec[,,lower.bound ~ peak.wl]@wavelength, xout = c(spec[,,peak.wl]$spc/2))

  upper <- approx(spec[,,peak.wl ~ upper.bound]$spc, spec[,,peak.wl ~ upper.bound]@wavelength, xout = c(spec[,,peak.wl]$spc/2))
 
  fwhm <- upper$y - lower$y
  center <- mean(c(upper$y,lower$y))
  skew <- 1 - (upper$y - peak.wl)/(fwhm/2)
  
  return(c(fwhm,center,skew))
}
```

Let's try it

```{r}
#Start a data frame to hold the data 

ex320data <- read.csv("V:/vuv-data/proj/PETPilot/Key/petpilot-key.csv")

ex320data$Em.Peak <- NA
ex320data$Em.Peak.int <- NA
ex320data$Em.Peak.fwhm <- NA
ex320data$Em.Peak.center <- NA
ex320data$Em.Peak.skew <- NA
ex320data$Em.Peak.skew.mag <- NA


# We'll use a for loop to run each spectrum in the hyperSpec Object 
row <- 1

for (i in 1:length(spec)) {
  
  # Define temporary holder for spectrum of interest
  
  temp <- spec[i]
  
  #Define the upper and lower bounds for the approx function
  
  lower.bound <- min(temp@wavelength)
    upper.bound <- max(temp@wavelength)
    
    # Find the maximum value in the region of interest
    peak <- max(temp[,,450~470])
    ex320data$Em.Peak.int[row] <- peak
    
    # Find the peak wavelength
    
    peak.out <- approx(temp[,,450 ~ 470]$spc, temp[,,450 ~ 470]@wavelength, xout = max(temp[,,450~470]))
    peak.wl <- ifelse(is.na(peak.out$y)==TRUE,450,peak.out$y)
    ex320data$Em.Peak[row] <- peak.wl
   
     # Run our function
    
    spec.data <- fwhm.spc(temp, as.numeric(peak.wl), lower.bound, upper.bound)
   
     # Store values in data frame
    
  ex320data$Em.Peak.fwhm[row] <- spec.data[1]
  ex320data$Em.Peak.center[row] <- spec.data[2]
  ex320data$Em.Peak.skew[row] <- spec.data[3]
  row <- row + 1
}
```

Check the result

```{r}
head(ex320data)
```

# Other plots

## Matrix Plot - Fluorescence Data
```{r}
plotmat(spec)
plotmat(spec, contour = TRUE, add= TRUE)
```

## Matrix Plot - IR Data
```{r}
plotmat(ir.spc)
```

## Stacked Spectra

```{r}
plotspc(spec, stacked = TRUE)
```

# Summary

## hyperSpec

- hyperSpec is a powerful R package for handling and manipulating spectral data in R
- More to come!

