---
title: "DSCI351-451- ISLR2,3 Intro R, Lin. Regr. Labs"
author: "Roger H. French, JiQi Liu"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---
 
 \setcounter{section}{11}
 \setcounter{subsection}{2}
 \setcounter{subsubsection}{2}

#### ISLR R Labs are at the end of each chapter


#### Chapter 2 Lab: Introduction to R

Notice: ISLR was published in 2013

ggplot2 Graphics are not used!

  - Since this is before ggplot2 (or even ggplot) were released
  
And there are no pipes!

  - So this was before the tidyverse R packages were released
  - not even the dplyr or plyr or magrittr R packages had been released
  
Also in ISLR, they use = instead of the proper <- assignment operator

  - So their code style is different from our Google R style

##### Basic Commands

```{r}
x <- c(1, 3, 2, 5) # concatenate a vector
x # dispay the vector
```

```{r}
x <-  c(1, 6, 2) # concatenate a vector
x # display a vector
y <-  c(1, 4, 3) # concatenate a vector

length(x) # get the length of a vector
length(y) # get the length of a vector

x + y # add two vectors
```

```{r}
ls() # list objects in the work space
rm(x, y) # remove objects in the work space
ls() # list objects in the workspace
rm(list = ls()) # remove all objects in the workspace 
```

```{r}
# ?matrix # help on matrix
x <- matrix(data = c(1, 2, 3, 4), nrow = 2, ncol = 2) # make a matrix
x # display a matrix
x <- matrix(c(1, 2, 3, 4), 2, 2) # make a matrix
matrix(c(1, 2, 3, 4), 2, 2, byrow = TRUE) # matrix filled by row
sqrt(x) # Square root of each value in the matrix
x^2 # Square each value in the matrix
```

```{r}
x <- rnorm(50) # randomly generates 50 values
y <- x + rnorm(50, mean = 50, sd = 0.1) # randomly generates 50 values with 
                    # certain criteria and adds the contents of the vector x

cor(x, y) # Finds the correlation between the two vectors
```

```{r}
set.seed(1303) # sets a new seed
rnorm(50) # makes 50 new random values

set.seed(3) # sets a new seed 
y <- rnorm(100) # Makes 100 random variables
```

```{r}
mean(y) # Calculates the mean of the vector
var(y) # calculates the variance of the vector
sqrt(var(y)) # calculates the square root of the variance
sd(y) # calculates the standard deviation. 
      # Which is the square root of the variance
```

##### Graphics

```{r}
x <- rnorm(100) # caluculates 100 random values
y <- rnorm(100) # calculates 100 random values

plot(x, y) # plots these two vectors against each other
```

```{r}
plot(x, y, xlab = "this is the x-axis", ylab = "this is the y-axis", 
     main = "Plot of X vs Y") # Plots the same data but applies labels to the plot
pdf("Figure.pdf") # starts the pdf device

plot(x, y, col = "green") # plots the data 
dev.off() # turns off the device and prints the pdf to file
```

```{r}
x <- seq(1, 10) # generates a sequence of 10 values
x # displays the above values
x <- 1:10 # this does the exact same thing as the sequence function (for integers) 
x # displays the vector
x <- seq(-pi, pi, length = 50) # Takes the two end limits and generates 50 numbers that fit inbetween
y <- x # sets x to y
f <- outer(x, y, function(x, y) cos(y)/(1 + x^2)) #x dotted with the transpose of y
contour(x, y, f) # generates a contour
contour(x, y, f, nlevels = 45, add = T) # Generates another contour plot with more contour lines
fa <- (f - t(f))/2 # a manipulation of the data
contour(x, y, fa, nlevels = 15) # Plots the new contour
image(x, y, fa) # turns the contour into a heatmap image
persp(x, y, fa) # creates a perspective plot of the data
persp(x, y, fa, theta = 30) # rotates the perspective plot by 30 degrees
persp(x, y, fa, theta = 30, phi = 20) # another rotation
persp(x, y, fa, theta = 30, phi = 70) # another rotation
persp(x, y, fa, theta = 30, phi = 40) # another rotation
```

##### Indexing Data

```{r}
A <- matrix(1:16, 4, 4) # generates a 4 by 4 matrix
A # displays the matrix
A[2, 3] # gets one value of the matrix 
A[c(1, 3), c(2, 4)] # grabs a few points from the matrix
A[1:3, 2:4] # gets a range of points from the matrix
A[1:2, ] # gets two rows
A[, 1:2] # gets two columns
A[1, ] # gets the first row
A[-c(1, 3), ] # gets not rows 1 and 3
A[-c(1, 3), -c(1, 3, 4)] # gets not those values
dim(A) # gets the dimensions of A
```

##### Loading Data

```{r}
library(ISLR)
attach(Auto)
# Auto <- read.table("Auto.data")
# fix(Auto)
# Auto <- read.table("Auto.data",header = T,na.strings = "?")
# fix(Auto)

Auto <- read.csv("data/Auto.csv", header = T, na.strings = "?")
fix(Auto)
dim(Auto)
Auto[1:4, ]
Auto = na.omit(Auto)
dim(Auto)
names(Auto)
attach(Auto)
```

##### Additional Graphical and Numerical Summaries

```{r}
library(ISLR)
plot(cylinders, mpg)
# plot(Auto$cylinders, Auto$mpg)
attach(Auto)
plot(cylinders, mpg)
cylinders = as.factor(cylinders)
plot(cylinders, mpg)
plot(cylinders, mpg, col = "red")
plot(cylinders, mpg, col = "red", varwidth = T)
plot(cylinders, mpg, col = "red", varwidth = T,horizontal = T)
plot(cylinders, mpg, col = "red", varwidth = T, xlab = "cylinders", ylab = "MPG")
hist(mpg)
hist(mpg, col = 2)
hist(mpg, col = 2, breaks = 15)
```

```{r}
# pairs(Auto)
pairs(~ mpg + displacement + horsepower + weight + acceleration, Auto)
plot(horsepower, mpg)
identify(horsepower, mpg, name)
summary(Auto)
summary(mpg)
```


#### Chapter 3 Lab: Linear Regression

```{r}
library(MASS)
library(ISLR)
```

##### Simple Linear Regression

```{r}
# fix(Boston)
names(Boston)

# lm.fit=lm(medv~lstat) # linear fit (fails)
lm.fit <- lm(medv~lstat, data = Boston) # linear fit (success with data)
attach(Boston) # attaches the dataset
lm.fit <- lm(medv~lstat) # linear model again
lm.fit # shows the lm fit

summary(lm.fit) # summary stats of the lm fit

names(lm.fit) # names of the components in the lm object

coef(lm.fit) # coefs of the model

confint(lm.fit) # 95% interval

predict(lm.fit, data.frame(lstat = (c(5,10,15))), interval = "confidence") # prediction with interval
predict(lm.fit, data.frame(lstat = (c(5,10,15))), interval = "prediction") # prediction with prediction interval

plot(lstat, medv) # plots the vars


abline(lm.fit) # plots the fit


abline(lm.fit, lwd = 3) # plots the fit again

abline(lm.fit, lwd = 3, col = "red") # plots the fit and adds a color

plot(lstat, medv, col = "red") # plots the data with a color

plot(lstat, medv, pch = 20) # plots the data with larger points

plot(lstat, medv, pch = "+") # plots the data with plus signs

plot(1:20, 1:20, pch = 1:20) # plots a x = y line of various shapes

par(mfrow = c(2, 2)) # sets of 2 by 2 plot window

plot(lm.fit) # plots various model plots. Residuals etc.

plot(predict(lm.fit), residuals(lm.fit)) # computes the residuals

plot(predict(lm.fit), rstudent(lm.fit)) # computes the student

plot(hatvalues(lm.fit)) # shows the leverage statistics

which.max(hatvalues(lm.fit)) # identifies the largest leverage statistic
```

##### Multiple Linear Regression

```{r}
lm.fit <- lm(medv~lstat + age, data = Boston) # plus sign adds more predictors
summary(lm.fit) # summary of the fit

lm.fit <- lm(medv~., data = Boston) # all predictors in data set used
summary(lm.fit) # summary of that model

library(car) # gets the car set which contains vif() 
vif(lm.fit) # compute variance inflation factors

lm.fit1 <- lm(medv~. - age, data = Boston) # remove age because of high p value

summary(lm.fit1) # summary of that model

lm.fit1 <- update(lm.fit, ~.-age) # updates the old model as opposed to making a new one
```

##### Interaction Terms

```{r}
# lstat:age adds an interaction term 
# however, the * adds the interaction terms as well as the individual terms
summary(lm(medv~lstat * age, data = Boston)) # summary of the interation terms and the individual terms
```

##### Non-linear Transformations of the Predictors

```{r}
lm.fit2 <- lm(medv~lstat + I(lstat^2)) # polynomial fit
summary(lm.fit2) # summary of that fit
lm.fit <- lm(medv~lstat) # a linear fit
anova(lm.fit, lm.fit2) # analysis of variance of those 2 fits
par(mfrow = c(2, 2)) # a 2 by 2 plot window
plot(lm.fit2) # plots the model plots into the 2 by 2 windows
lm.fit5 <- lm(medv~poly(lstat, 5)) # computes a fifth order polynomial
summary(lm.fit5) # summary of the fifth order poly
summary(lm(medv~log(rm), data = Boston)) # summary of the log of the rm predictor 
```

##### Qualitative Predictors

```{r}
# fix(Carseats)
names(Carseats)
lm.fit <- lm(Sales~. + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit)
attach(Carseats)
contrasts(ShelveLoc)
```

##### Writing Functions

As we have seen, R comes with many useful functions, and still more func-
tions are available by way of R libraries. 

However, we will often be interested in performing an operation for which 
no function is available. 

In this setting, we may want to write our own function. 

For instance, below we provide a simple function 

  - that reads in the ISLR and MASS libraries, 
  - called LoadLibraries(). 

Before we have created the function, R returns an error if we try to call it.

```{r}
# LoadLibraries
# LoadLibraries()
```

We now create the function. 

Note that 

  - the + symbols are printed by R 
  - and should not be typed in. 
  
The { symbol informs R that multiple commands are about to be input. 

Hitting Enter after typing { 

  - will cause R to print the + symbol. 
  
We can then input as many commands as we wish, hitting Enter after each one. 

Finally the } symbol informs R that no further commands will be entered.

```{r}
LoadLibraries = function(){
  library(ISLR)
  library(MASS)
  print("The libraries have been loaded.")
  }
```

Now if we type in LoadLibraries, R will tell us what is in the function.

```{r}
LoadLibraries
```

If we call the function, 

  - the libraries are loaded in 
  - and the print statement is output.

```{r}
LoadLibraries()
```

