---
title: "Basics of R"
author: "Alan Curran, Roger French"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

 \setcounter{section}{1}
 \setcounter{subsection}{1}
 \setcounter{subsubsection}{2}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 6, fig.height = 4.5) 
```

### Basics of R

#### Math operations 

```{r}
  2 + 2
  8 / 4
  2 * 3
  3 ^ 3
  a = 3
  b = 3
  a + b
```

  - Assignment operator should ALWAYS be used instead of =
  - It is directional
  - And it meets Google's R style guide
    - You can findd the style guide in 3-readings>3-CheatSheets
  - = works in most cases too but may cause problems

```{r}
  a <- 3
  b <- 6
  
  a <- b
  a
  
  a <- 3
  b <- 6
  
  a -> b
  a
```

#### Object Classes

```{r}
  # numerics
  a <- 5
  class(a)
  # integers
  b <- as.integer(42)
  class(b)
  # logicals
  c <- TRUE
  class(b)
  # characters
  d <- "hello world"
  class(c)
  # factors
  e <- as.factor(c("1", "1", "a", "1", "c", "a"))
  class(e)
  e
  # quick note it you want to convert a factor to a numeric 
  # You have to convert it to a character first, then a numeric
```

#### Indexing

  - R uses brackets to reference a data index
  - data["row","column"]
  - Standard organization for data set has varaibles as columns and observations as rows
  - Keep in mind that R indexing starts from 1, not 0
  - We can load a test data set for indexing

```{r}
  data("iris")
  
  iris[1,2]
  
  # Leaving a row or column input blank puts all values
  # First column
  iris[,1]
  
  # First row
  iris[1,]
  
  # Data frames have associated column names
  colnames(iris)
  
  # Columns can be called by name using $
  # Rstudio features tab completion for thing like column names
  iris$Sepal.Length
```

#### Built-in Functions

  - Functions are processes that take an input and give an output
  - Rstudio has tab completion for function inputs

```{r}
  max(iris$Petal.Length)
  mean(iris$Sepal.Width)
  sd(iris$Petal.Width)
  
  # Functions can take multiple inputs, they can be named in the call or placed in order
  plot(x = iris$Sepal.Length, y = iris$Sepal.Width)
  
  # x and y can be specified with x = ... in any order or the inputs can be given in order
  # This plot is the same as the previous
  plot(iris$Sepal.Length, iris$Sepal.Width)
```

#### Matrix operations

```{r}
  mat <- matrix(data = 1:9, nrow = 3, ncol = 3)
  mat
  
  # Element multiplication
  mat*mat
  
  # Matrix multiplication
  mat %*% mat
  
  # t() function is for transposing
  t(mat)
  mat %*% t(mat)
  
  # Inverse matrix
  mat[2,3] <- 18
  solve(mat)
  solve(mat) %*% mat
```

#### Structures in R

```{r}
  # for loops
  for (i in 1:5) {
    print(i)
  }
  
  # While loops 
  i <- 10
  while (i > 5) {
    i <- i - 1
    print(i)
  }
  
  # if statments
  dave <- TRUE
  # if (dave) {} also works
  if (dave == TRUE) {
    print("good morning dave")
  }
  
  # User defined functions
  math <- function(a,b) {
    c <- a + b*2
    # return defines what the output of the function is
    return(c)
  }
  math(2,6)
```



### Recommended Functions

This is a list of functions in base R I would recommend knowing. 

  - There are countless more functions and packages to use, 
    - but these are cover most of the basics of data manipulation. 
  - Additional information can be found by entering ?<function name> in the R console.

#### Useful R Functions

  - install.packages()
  - library()
  - read.csv()
  - readlines()
  - class()
  - min()
  - max()
  - median()
  - mean()
  - sd()
  - c()
  - rbind()
  - merge()
  - paste()
  - ifelse()
  - subset()
  - complete.cases()
  - is.na()
  - which()
  - grep()
  - grepl()
  - gsub()
  - as.numeric()
  - as.character()
  - as.factor()
  - list.files()
  - plot()
  - abline()
  - points()
  - lm()
  - summary()


