---
title: "CWRU DSCI351-351M-453: Week10b w10b-p-LinRegr"
author: "Roger H. French, JiQi Liu"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
bibliography: refs.bib
---
  
\setcounter{section}{10}
\setcounter{subsection}{2}
\setcounter{subsubsection}{1}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 6, fig.height = 4.5) 
```

<!-- 
  How to make comments inside Rmarkdown
# Script Name: My class notes template for Fall 2018
# Purpose: This is a template Rmd file to start a new class from
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike NonCommercial 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - Filename.Rmd - Roger French started this blank Rmd script
-->
  
  <!-- Or on a single line like this -->

#### Understanding simple linear regression
 
##### Build and use our own simple linear regression algorithm

  - Create multiple linear regression models in R
  - Perform diagnostic tests of such models
  - Score new data using a linear regression model
  - Examine how well the model predicts the new data

Regression seeks to obtain the model coefficients

  - that explain the variable's relationship the best 
  - but such a model only seldom reflects the relationship entirely

Indeed, measurement error, 

  - And also attributes that are not included in the analysis 
  - affect also the data. 

The model residuals 

  - express the deviation of the observed data points 
  - to the model. 

The residual's value 

  - is the vertical distance from a point 
  - to the regression line. 

#### Let's examine this with an example of the iris dataset. 

We have already seen that the dataset contains data about iris flowers. 

For the purpose of this example, 

  - we will consider the petal length as the response
    - sometimes the response is referred to as the "criterion"
  - and the petal width as the predictor

```{r}
plot(iris$Petal.Length ~ iris$Petal.Width, 
     main = "Relationship between petal length and petal width", 
     xlab = "Petal width", ylab = "Petal length")
iris.lm = lm(iris$Petal.Length ~ iris$Petal.Width)
abline(iris.lm)
```

##### Computing the intercept and slope coefficient

```{r}
SlopeCoef = cor(iris$Petal.Length,iris$Petal.Width) * 
  (sd(iris$Petal.Length) / sd(iris$Petal.Width))
SlopeCoef

coeffs = function(y,x) {
  ((length(y) * sum( y*x)) - 
      (sum( y) * sum(x)) )  / 
    (length(y) * sum(x^2) - sum(x)^2)
}

coeffs(iris$Petal.Length, iris$Petal.Width)
```

##### Now make your linear regression function
```{r}
iris.lm

regress =  function(y,x) {
  slope = coeffs(y,x)
  intercept = mean(y) - (slope * mean(x))
  model = c(intercept, slope)
  names(model) = c("intercept", "slope")
  model
}
```

##### Now perform regression on Petal Length and Petal Width
```{r}
model = regress(iris$Petal.Length, iris$Petal.Width)
model
```

##### Obtaining the residuals
```{r}
resids = function(y,x, model) {
  y - model[1] - (model[2] * x)
}

Residuals = resids(iris$Petal.Length, iris$Petal.Width, model)

head(round(Residuals,2))

par(mfrow = c(2, 2))
plot(iris.lm)
```

#### Computing the significance of the coefficients

This is also the uncertainty 

  - in your regression coefficients
```{r}
Significance = function(y, x, model) {
  SSE = sum(resids(y,x,model)^2)
  DF = length(y) - 2
  S = sqrt( SSE / DF)
  SEslope = S / sqrt(sum( (x - mean(x))^2 ))
  tslope = model[2] / SEslope
  sigslope = 2*(1 - pt(abs(tslope),DF))
  SEintercept = S * sqrt((1/length(y) + mean(x)^2 / sum( (x - mean(x))^2)))
  tintercept = model[1] / SEintercept
  sigintercept = 2*(1 - pt(abs(tintercept),DF))
  RES = c(SEslope, tslope, sigslope, SEintercept, tintercept, sigintercept)
  names(RES) = c("SE slope", "T slope", "sig slope", "SE intercept", 
                 "t intercept", "sig intercept")
  RES
} 

round(Significance(iris$Petal.Length,iris$Petal.Width, model), 3)

summary(iris.lm)
```


#### Links

Learning Predictive Analytics with R, Eric Mayor, Packtpub 2015