---
title: 'CWRU DSCI351-451: Semester Project 2- Time Series Analysis'
author: "Roger H. French, JiQi Liu"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 6
    highlight: tango
  html_document:
    toc: yes
urlcolor: blue
---

<!--
# Script Name: 1710-dsci351-451-project2.Rmd
# Authors: Roger H. French
# License: Creative Commons Attribution-ShareAlike 4.0 International License.
##########
# Latest Changelog Entires:
# v0.00.01 - 1501cwru-dsci-NAMEIT.RMD - Roger French started this blank Rmd
##########

# Rmd code goes below the comment marker!
-->


\setcounter{section}{2}
\setcounter{subsection}{1}
\setcounter{subsubsection}{0}

# Semester Project 2: Time Series Analysis

Time series are a common type of data, 

  - consisting of measurements that are continuous over a time range. 

In this project we will be using classical decomposition 

  - to perform analysis on a time series.

First as an introduciton to decomposition we will have a quick example.

## Problem 1

  - What is the decomposition of a timeseries?
  - The AirPassangers data set of airline passangers every month for 12 years

```{r}
library(datasets)
air <- as.data.frame(AirPassengers)
```

  - Plot the total time series of air passangers 
  - What do you notice about the plot?

```{r}

```

  - Use the ts() function in base R 
    - to define AirPassangers as a time series with a yearly trend 
  - If the data is taken monthly, 
    - what will the frequency (points per season) of a yearly season be? 

```{r}

?ts()

```

  - Use the decompose() function 
    -to demcompose the time series and remove the seasonality
  - The type for this time series is multiplicative
  - Plot the decomposed time series, what do you notice about the trend?

```{r}


```

  - Isolate the trend and plot the trend on top of the raw data with the seasonality included
  - How well does the trend represent the data?

```{r}

```

## Problem 2

Now lets try this with a real world time series. 
We'll be using one month of power and weather data from a solar power plant.
The data set variables are as follows:

  - time: The timestamp at which the data was taken
  - ghir: Global Horizontal Irradiance from a sensor at the site, 
    - the power from the sunlight over an area normal to the surface of the earth $(Watts/m^2)$
  - iacp: The AC power from the power plant $(kW)$
  - temp: The air temperature $(Celsius)$
  - ghi_solargis: The Global Horizontal Irradiance, not from a sensor, 
    - but predicted using weather modeling $(Watts/m^2)$
  - clear: A logical indicating whether the sky was "clear" during measurement, 
    - determined by comparing the ghi and ghi_solargis data
  - ratio: the ratio of the Global Horizontal Irradiance 
    - and the Plane of Array Irradiance (the irradiance normal to the surface of a tilted module)

The power from solar panels is dependant on the irradiance hitting it, 

  - so a power reading is often meaningless without a corresponding irradiance measurement.

It is useful to have multiple sources of irradiance measurements. 

Sensors on the ground are useful because 

  - they strongly represent the irradiance that is hitting the module; 
  - however, sensors can begin to drift if not cleaned or calibrated properly. 
  - An unstable sensor can render an entire data set useless.

To combat this, we also have irradiance data from SolarGIS, 

  - a company that uses satellite images to model and predict 
  - the irradiance at the surface of the earth. 

  - Plot the irradiance and power for the first week of data, 
    - how does the irradiance look compared to 
      - what you would expect from the trend of sunlight? 
    - How well does the power and irradiance match up?

```{r}

```

- Use the ts() functions and the stlplus() function from the stlplus package 

  - to decompse the sensor and SolarGIS irradiance and the power 
    - for the whole month. 
  -  Plot each of the decompositions, what do you notice?

```{r}
# think carefully about the frequency you'll need to define for this data
# what is the seasonal component to this data and how nay data points make up a season?
# use s.window = "periodic" for the stlplus function
library(stlplus)
?stlplus()

```

  - Isolate the trends for the 3 time series you just decomposed 
    - and build a linear model for each one.
  - Compare the models to each other, how are they different?

```{r}

```

  - Solar panel degradation leads to less power output over time 
    - at the same irradiance conditions.
  - Based on the linear models you found for the trends of power and irradiance, 
    - is this system degrading over time?
  - How do the sensor GHI and the SolarGIS GHI compare to power?





#### Links
 
http://www.r-project.org 

http://rmarkdown.rstudio.com/  

<!--
# Keep a complete change log history at bottom of file.
# Complete Change Log History
# v0.00.00 - 1405-07 - Nick Wheeler made the blank script
##########

